Nearest Neighbour Methods for Length of Stay Prediction

### 2015-02-07 papers/paper.tex
Small rearrangements, fleshing out the evaluation section

Like the Na\"{i}ve Bayes classifier, $k$-NN allows all features to be taken
into account equally when classifying a new example. However, since feature
values often span different ranges and units of measurement, a particularly
large range of values for a feature would contribute much more to the distance
than another feature that spans a lower set of values.
To avoid this, it is important to \textit{normalise} the
values of each numeric feature before using $k$-NN: for each feature $i$, we
calculate the normalised value of the feature for each training example. This
is done using the following relationship:
\begin{equation*}
a_i = s\left(\dfrac{v_i - \mathrm{min }v_i}{\mathrm{max }v_i - \mathrm{min }v_i}\right) + t
\end{equation*}
where $a_i$ is the normalised feature value, $v_i$ is the actual value in the
data set, max $v_i$ and min $v_i$ are taken over all training examples, $s$ is
a scaling factor and $t$ indicates how much to translate the range by. If $s=1$
and $t=0$, then the range of normalised values will be in $[0,1]$.

@@

\section{Evaluation}
\label{sec:eval}
\subsection{Data Set}
Our evaluation was carried out using a trauma-specific data set
containing 2546 records collected from 2007--2011 collected at a major hospital
in Sydney, Australia.

\subsection{Discretisation}
Fayyad
and Irani's entropy-based method of discretisation \cite{Fayyad1993}:
to discretise an attribute,
each possible split point is considered and the information gain from that
split is computed. The split point with the highest information gain is
therefore chosen. Once the first split is determined, the algorithm continues
recursively in the newly created upper and lower ranges.

The \textit{minimum description length (MDL)} principle is used to determine
the stopping criterion. The MDL value is calculated by adding the information
values from two cases: split or do not split. If we split, then we can encode
the split point in $\mathrm{log}_2(N-1)$ bits (where $N$ is the number of
instances) and subsequently the classes of those above and below it. If we do not split,
then each instance's label must be encoded. Mathematically, given $N$ examples
with entropy $E$, sub-intervals $k_1$ and $k_2$ with the entropy of examples
in each sub-interval denoted $E_1$ and $E_2$:
\begin{equation*}
\mathrm{MDL} = \dfrac{\mathrm{log}_2(N-1)}{N} + \dfrac{\mathrm{log}_2(3^k-2) - kE + k_1E_1 + k_2E_2}{N}
\end{equation*}
The first term is the information needed for the splitting point, and the
second is a correction necessary for encoding which classes belong to which
sub-interval. The decision rule is then to split if the information gain for
that split exceeds the MDL value.

Since this requires exhaustive searching of the full range of feature values
to find potential split points, this method of supervised discretisation can
be time-consuming given large enough data sets. However, it has been proven
that any potential split that minimises the information value will never occur
between two examples of the same class \cite{Witten2005}.
This means that we only need to
consider splits between regions of different classes, reducing the number of
possible points to consider.

The data was discretised according to the above method using the supervised
discretisation filter in WEKA. We kept both the original data set (without
discretisation) and the discretised data set for comparison.

\subsection{Classification Algorithms}
We tested the performance of Zero-R, 1R, Na\"{i}ve Bayes, C4.5 decision trees,
support vector machines, multi-layer perceptrons, $k$-nearest neighbour with 1
and 20 nearest neighbours, K* and logistic regression, as well as our proposed
Ranked Distance Nearest Neighbour.

\subsection{Experimental Setup and Procedure}
The steps that we took to evaluate the performance of the various classifiers
and feature selection methods were:
\begin{enumerate}
\item For each data set, we evaluated the effect of the 4 automatic feature
selection methods on each of the 8 classifiers listed above, giving us 32
configurations for each data set, over
4 sets. Evaluation was performed using ten runs of \textit{ten-fold stratified
cross-validation}.
\item For the two trauma LOS data sets, we additionally evaluated the two
manual feature selection methods with all 8 classifiers. Recall from Section
\ref{sec:manual} that one of the feature sets is from the key previous work
on trauma LOS prediction that we use as a baseline, and the other feature set
is a set of features included through consultation with a domain expert.
\end{enumerate}

@@

\begin{subfigure}{.48\textwidth}
\includegraphics[width=\textwidth]{../thesis/images/results/tr-cfs-acc.eps}
\caption{}
\label{fig:tr-nothreshold-cfs-acc}
\end{subfigure}%
\begin{subfigure}{.55\textwidth}
\includegraphics[width=\textwidth]{../thesis/images/results/tr-cfs-auc.eps}
\caption{}
\label{}
\end{subfigure}
\caption[]{Comparison of accuracy and AUC between classifiers for each non-threshold feature selection method, grouped by whether or not discretisation was applied.}
\label{fig:tr-nothreshold}
\end{figure*}

\begin{figure*}
\ContinuedFloat

>>>>

Like the Na\"{i}ve Bayes classifier, $k$-NN allows all features to be taken
into account equally when classifying a new example. However, since feature
values often span different ranges and units of measurement, a particularly
large range of values for a feature would contribute much more to the distance
than another feature that spans a lower set of values.
To avoid this, it is important to \textit{normalise} the
values of each numeric feature before using $k$-NN: for each feature $i$, we
calculate the normalised value of the feature for each training example. This
is done using the following relationship:
\begin{equation}
\label{eq:normalise}
a_i = s\left(\dfrac{v_i - \mathrm{min }v_i}{\mathrm{max }v_i - \mathrm{min }v_i}\right) + t
\end{equation}
where $a_i$ is the normalised feature value, $v_i$ is the actual value in the
data set, max $v_i$ and min $v_i$ are taken over all training examples, $s$ is
a scaling factor and $t$ indicates how much to translate the range by. If $s=1$
and $t=0$, then the range of normalised values will be in $[0,1]$.

@@

\section{Evaluation}
\label{sec:eval}
\subsection{Data Set}
The data set used consisted of trauma registry data from the trauma centre
at the Royal Prince Alfred Hospital, a major trauma centre in New South Wales,
Australia. It covered all adult (age 15 and over) inpatient admissions to the
trauma centre from 2007--2011. 
All patients were first admitted to the trauma ward until discharged
or transferred to an appropriate unit within the hospital. A single trained
data manager recorded a variety of attributes about the admitted patient,
such as age, gender, blood pressure, mechanism of injury and body regions
that were injured.

There were 2546 patient records in the data set we received, comprising of 79
features, one of which was the target variable \texttt{los48}.
\texttt{los48} was a binary variable -- that is, it could only take two
values, 0 or 1 -- with 1 indicating that the patient stayed two days or less,
and 0 indicating a stay of longer than 2 days.

\subsection{Preprocessing}
The data set was cleaned by correcting spelling and punctuation
inconsistencies and removing extraneous whitespace. Examples which had missing
values for at least one feature were removed, in line with previous work on the
same data set.

Continuous-valued features were then normalised into the $[0,1]$ range as
described in Section \ref{sec:nn} above, substituting $s=2$ and $t=-1$ into
Equation \ref{eq:normalise}. These continuous features were then discretised
using Fayyad and Irani's method of supervised discretisation
\citep{Fayyad1993}. We kept both the original data set (normalised, without
discretisation) and the discretised (and normalised) data set for comparison.

\subsection{Classification Algorithms}
We tested the performance of
support vector machines, multi-layer perceptrons, $k$-nearest neighbour with 1
and 20 nearest neighbours, K* and logistic regression, as well as our proposed
Ranked Distance Nearest Neighbour -- a total of 7 different classifiers.

\subsection{Statistical Significance}

\subsection{Experimental Setup and Procedure}
The steps that we took to evaluate the performance of the various classifiers
and feature selection methods were:
\begin{enumerate}
\item For both the discretised and non-discretised data set, we trained all 7
classifiers without feature selection. Evaluation was performed using ten runs
of \textit{ten-fold stratified cross-validation}.
\item Then, for each data set, we evaluated the effect of the 3 automatic
feature
selection methods on each of the 7 classifiers listed above, giving us 21
configurations for the discretised and non-discretised data set.
\item In order to ensure that the baseline in the literature was comparable
to our work, we used the baseline features and applied their method, logistic
regression, to obtain a measure of performance on the non-discretised data set.
We also evaluated this feature set against the rest of the 6 classifiers, and
repeated this using the discretised data.
\item Additionally, the expert-selected features were tested against all 7
classifiers, with both the discretised and non-discretised data.
\end{enumerate}

@@
